---
title: "BDA: Chapter 10 exercices"
author: "OS"
output: html_document
---

```{r, include = FALSE}
source("R/setup.R")
```

# Exercice 1

## a

Not sure if this can count as "mathematically", but deriving the sd of a quantile
of the normal distribution doesn't seem to be particularly easy.

Let $\{x_i\}_{i \in 1,\dots,n}$ be an $n$ sample of $\theta$. We are intrested in
bounding the sd of its $p$th, that we write $\theta_{(\lceil np\rceil)}$. Note
that the number observations that falls above the $p$th quantile follows a
$\text{Binom}(n, p)$ distribution. We use this fact to estimate the sd.

The mean and sd of this distribution is $np$ and $\sqrt{np(1-p)}$ respectively.
We estimate the sd of the quantile as $\frac{1}{2}(
\theta_{(\text{round}(np + \sqrt{np(1-p)}))} -
\theta_{(\text{round}(np - \sqrt{np(0-p)}))}
)$.


An interesting note, I first took the floor and ceiling for the lower and upper
quantile respetively, this however gave biased results. The calculated sd was
too large. Rounding appers to give much better results.

Another way to do this is to use the (known but ugly) distribution of each
quantile, and determine its sd.

## b

We check the above result via simulations. First we implemented the above
procedure.

```{r}
quantile_sd <- function(x, p) {
    n <- length(x)
    e_b <- binom <- n * p
    sd_b <- sqrt(n * p * (1 - p))
    up <- round(e_b + sd_b)
    lo <- round(e_b - sd_b)
    ord <- sort(x, decreasing = FALSE)
    0.5 * (ord[up] - ord[lo])
}

# test, reasonable enough
quantile_sd(runif(1000), 0.5)
quantile_sd(runif(10), 0.5)
quantile_sd(runif(10), 0.9)
```


To test how well this works, we calculate the real sd via simulations.

```{r}
p <- 0.975
n <- 100
sd(replicate(1e3, quantile(rnorm(n, 0, 1), p)))
mean(replicate(1e3, quantile_sd(rnorm(n, 0, 1), p)))
```

We can even do better by comparing the sd over various sample sizes.

```{r}
n <- c(1:100, seq(from = 100, to = 10e3, by = 100))
true_sd <- sapply(
    n,
    \(n) sd(replicate(100, quantile(rnorm(n, 0, 1), p)))
)
estim_sd <- sapply(
    n,
    \(n) mean(replicate(1e3, quantile_sd(rnorm(n, 0, 1), p)))
)

bind_rows(
    true = enframe(true_sd),
    estim = enframe(estim_sd),
    .id = "type"
) %>%
    mutate(n = rep(n, 2)) %>%
    ggplot() +
    aes(
        x = n,
        y = value,
        color = type,
        group = type
    ) +
    geom_line() +
    geom_hline(yintercept = 0.1 * 1)
```

This pretty much allows 

# Checking Vim

A mistake in a sentense.

```{lua}
vim.lsp.start({
   name = 'My R LSP',
   cmd = {'R.exe', '--slave', '-e', 'languageserver::run()'},
   root_dir = 'C:/Documents/Projects/bayes',
})
```



